Generate data for NOBA cod ensemble modeling

### 2020-03-20 NOBAcod.Rmd
test loading full age structure

```{r full-test}

@@

>>>>

```{r full-test, message=FALSE, warning=FALSE}

@@

What is in the ANNAGEBIO and ANNAGECATCH .nc files?

```{r annagebio-contents}

# taken from existing load_nc

file.nc <- file.path(d.name, "outputnordic_runresults_01ANNAGEBIO.nc")

  # Load ATLANTIS output!
  at_out <- RNetCDF::open.nc(con = file.nc)

  # Get info from netcdf file! (Filestructure and all variable names)
  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length

RNetCDF::close.nc(at_out)

ANNAGEBIO.nc variable names (first hundred): `r var_names_ncdf[1:100]`

ANNABEBIO.nc variable names (last hundred): `r var_names_ncdf[1590:1690]`

ANNAGEBIO.nc timesteps: `r n_timesteps`

ANNAGEBIO.nc boxes: `r n_boxes`

ANNAGEBIO.nc layers: `r n_layers`

Annual age output is in numbers and weight at each timestep, layer, and box, which is really good news. We should be able to get weight at true age more directly this way than the interpolation I was planning. Also, we may be able to do length estimation differently?

To modify `load_nc_annage` we need to figure out how many true ages each species has from the biological parameter file using `load_biolprm`:

```{r biolprm}

biolprm <- load_biolprm(d.name, biol.prm.file)
fgs <- load_fgs(d.name, functional.groups.file)

#max age should be ages per cohort times the number of age classes (<10 for some NOBA spp)

names(biolprm$agespercohort) <- c("code", "ageperagecl")

maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
    select(code, nagecl, ageperagecl) %>%
    mutate(maxage = nagecl * ageperagecl) %>%
    mutate(name = fgs$Name[match(code, fgs$Code)])

select_groups <- c('Mackerel', 'North_atl_cod', 'Capelin')

for (i in seq_along(select_groups)) {
  select_group_ages <- 1:maxage$maxage[match(select_groups[i], maxage$name)]
}

Try this `load_nc_annage` function:

```{r loadncannage, eval=FALSE}

load_nc_annage <- function(dir = getwd(), file_nc, bps, fgs, biolprm, select_groups,
  select_variable =
  c("Nums", "Weight"),
  check_acronyms = TRUE, bboxes = c(0), verbose = FALSE) {
  # NOTE: The extraction procedure may look a bit complex...
  # A different approach would be to
  # create a dataframe for each variable (e.g. GroupAge_Nums)
  # and combine all dataframes at the end.
  # This alternative approach requires a lot more storage
  # and the code wouldn't be vectorized.

  # Check input of the nc file
  if (tail(strsplit(file_nc, "\\.")[[1]], 1) != "nc") {
    stop("The argument for file_nc,", file_nc, "does not end in nc")
  }
  if (is.null(dir)) {
    file.nc <- file_nc
  } else {
    file.nc <- file.path(dir, file_nc)
  }

  # Check input of select_variable as only one value is allowed
  select_variable <- match.arg(select_variable, several.ok = FALSE)

  # Check input structure!
  if (check_acronyms) {
    active_groups <- as.vector(subset(fgs, IsTurnedOn == 1)$Name)
    inactive_groups <- select_groups[which(
      !is.element(select_groups, active_groups))]
    if (length(inactive_groups) >= 1) {
      select_groups <- select_groups[!is.element(select_groups, inactive_groups)]
      warning(paste(paste("Some selected groups are not active in the model run.",
        "Check 'IsTurnedOn' in fgs\n"),
        paste(inactive_groups, collapse = "\n")))
    }
    if (all(!is.element(select_groups, active_groups))) {
      stop(paste("None of the species selected are active in the model run.",
        "Check spelling and Check 'IsTurnedOn' in fgs"))
    }
  }

  # Deal with file structures

  # Load ATLANTIS output!
  at_out <- RNetCDF::open.nc(con = file.nc)
  on.exit(RNetCDF::close.nc(at_out))

  if (select_variable != "N" & all(is.element(select_groups, bps))) {
    stop("The only output for Biomasspools is N.")
  } else{
    print(paste("Read", file.nc, "successfully"))
  }

  # Get info from netcdf file! (Filestructure and all variable names)
  var_names_ncdf <- sapply(seq_len(RNetCDF::file.inq.nc(at_out)$nvars - 1),
    function(x) RNetCDF::var.inq.nc(at_out, x)$name)
  n_timesteps <- RNetCDF::dim.inq.nc(at_out, 0)$length
  n_boxes     <- RNetCDF::dim.inq.nc(at_out, 1)$length
  n_layers    <- RNetCDF::dim.inq.nc(at_out, 2)$length

  # Extract data from the ncdf file
  # Create a vector of all potential variable names
  # Only use names which are available in the ncdf-file as an
  # extraction of missing variables is not possible
  # Create vector of available species at the end using search_clean
  # This is needed to create species-names later on

  # CHANGED hardcoded cohort number to reflect true ages from

  # To make the creation of variables as robust as possible
  # we introduce different combinations of groups, variable, and cohort
  # Only combinations present in the ncdf are used later on
  # Loop over select_groups to use the same ordering

  names(biolprm$agespercohort) <- c("code", "ageperagecl")

  maxage <- merge(biolprm$maturityogive, biolprm$agespercohort, all.x = T) %>%
    select(code, nagecl, ageperagecl) %>%
    mutate(maxage = nagecl * ageperagecl) %>%
    mutate(name = fgs$Name[match(code, fgs$Code)])

  search <- list()
  for (i in seq_along(select_groups)) {
    select_group_ages <- 1:maxage$maxage[match(select_groups[i], maxage$name)]
    search[[i]] <- c(
      unlist(lapply(paste0(select_groups[i], select_group_ages),
                    paste0, select_variable)),           # GroupCohortVariable
      unlist(lapply(paste0(select_groups[i], select_variable),
                    paste0, select_group_ages)),                   # GroupVariableCohort
      unlist(lapply(paste0(select_groups[i], select_group_ages),
                    paste, select_variable, sep = "_")), # GroupCohort_Variable
      unlist(lapply(paste(select_groups[i], select_variable, sep = "_"),
                    paste0, select_group_ages)),                   # Group_VariableCohort
      unlist(lapply(paste(select_groups[i], select_group_ages, sep = "_"),
                    paste, select_variable, sep = "_")), # Group_Cohort_Variable
      unlist(lapply(paste(select_groups[i], select_variable, sep = "_"),
                    paste, select_group_ages, sep = "_")),         # Group_Variable_Cohort
      unlist(paste0(select_groups[i], select_variable)),                                                      # GroupVariable
      unlist(paste(select_groups[i], select_variable,
                   sep = "_"))                           # Group_Variable
      )
    search[[i]] <- search[[i]][is.element(search[[i]], var_names_ncdf)]
    search[[i]] <- unique(search[[i]])
  }
  search_clean <- do.call(c, search)
  # If the combination of select_groups and select_variable ends up not being found.
  if (length(search_clean) == 0) return(0)

  at_data <- lapply(search_clean, RNetCDF::var.get.nc, ncfile = at_out)

  # Get final species and number of ageclasses per species
  final_species <- select_groups[sapply(
    lapply(select_groups, grepl, x = search_clean), any)]
  final_agecl <- maxage$maxage[
    sapply(final_species, function(x) which(x == maxage$name))]

  num_layers <- RNetCDF::var.get.nc(ncfile = at_out, variable = "numlayers")[, 1]
  # add sediment layer!
  num_layers <- num_layers + ifelse(num_layers == 0, 0, 1)

  # Create an array of layerids.
  # Every entry in the array indicates if a layer is present (= 1) or not (= 0).
  # Boxes without layers (= islands) have only 0s as id,
  # used later on to remove data from non-existent layers!
  # By default output should be 0 in those layers.
  # Layers in boundary boxes are set to 0 if bboxes is anything other than NULL!
  # Applying a boolean array to an array results in a vector!
  for (i in seq_along(num_layers)) {
    if (i == 1) layerid <- array(dim = c(n_layers, n_boxes))
    if (num_layers[i] == 0) {
      layerid[, i] <- 0
    } else {
      if (!is.null(bboxes) & is.element((i - 1), bboxes)) {
        layerid[, i] <- 0
      } else {
        layerid[, i] <- c(rep(0, times = n_layers - num_layers[i]),
                          rep(1, times = num_layers[i]))
      }
    }
  }

  # Create vectors for polygons and layers
  # Each vector has the length equal to one time-step
  # All data from islands and non-existent layers is removed
  # Therefore the length of these
  # vectors is equal for each extracted variable
  boxes <- 0:(n_boxes - 1)
  # Remove islands and boundary boxes
  island_ids <- num_layers == 0
  if (!is.null(bboxes)) {
    boundary_ids <- is.element(boxes, bboxes)
    island_ids <- island_ids | boundary_ids
  }
  boxes <- boxes[!island_ids]
  num_layers <- num_layers[!island_ids]

  polygons <- rep(boxes, times = num_layers)
  layers <- sapply(num_layers[num_layers != 0] - 2,
    function(x) c(seq(x, from = 0, by = 1), n_layers - 1))
  if (any(sapply(layers, length) != num_layers[num_layers != 0])) {
    stop("Number of layers incorrect.")
  }
  layers <- do.call(c, layers)
  if (length(polygons) != length(layers)) {
    stop("Number of polygons and layers do not match.")
  }

  # In the following section the data is transformed to a long dataframe
  # I haven't found any solution to vectorize the creation of the dataframe
  # columns (species, age, polygons,...)
  # when data from 2d and 3d arrays
  # (e.g. select_variable = "N" all biomasspools are only present in the
  # sediment layer.) are read in simultaneously.
  # Therefore the current "messy" solution splits the data
  # in 2 subpopulations: 2d-data and 3d-data
  at_data3d <- at_data[which(sapply(at_data, function(x) length(dim(x))) == 3)]
  at_data2d <- at_data[which(sapply(at_data, function(x) length(dim(x))) == 2)]

  int_fs <- final_species
  int_fa <- final_agecl

  if (length(at_data3d) >= 1) {
    # Remove biomasspools if selected variable is "N"!
    if (select_variable == "N") {
      int_fs <- final_species[!is.element(final_species, bps)]
      int_fa <- final_agecl[!is.element(final_species, bps)]
      # Note this only works if age-structured vertebrates have 10 ageclasses
      int_fa[int_fa == 10] <- 1
    }
    for (i in seq_along(at_data3d)) {# for loop over all variables
      if (i == 1) result3d <- list()
      for (j in 1:n_timesteps) {# loop over timesteps
        if (j == 1) values <- array(dim = c(length(layers), n_timesteps))
        values[, j] <- at_data3d[[i]][,, j][layerid == 1]
      }
      result3d[[i]] <- as.vector(values)
    }
    result3d <- data.frame(
      species = unlist(sapply(
        X = mapply(FUN = rep, x = int_fs, each = int_fa, SIMPLIFY = FALSE,
        USE.NAMES = FALSE),
        FUN = rep, each = length(layers) * n_timesteps, simplify = FALSE)),
      agecl = unlist(sapply(
        X = sapply(X = int_fa, FUN = seq, from = 1, by = 1, simplify = FALSE,
        USE.NAMES = FALSE),
        FUN = rep, each = length(layers) * n_timesteps, simplify = FALSE)),
      polygon = unlist(sapply(
        X = n_timesteps * int_fa, FUN = rep, x = polygons, simplify = F,
        USE.NAMES = FALSE)),
      layer = unlist(sapply(
        X = n_timesteps * int_fa, FUN = rep, x = layers, simplify = FALSE,
        USE.NAMES = FALSE)),
      time = unlist(sapply(
        X = int_fa, FUN = rep, x = rep(0:(n_timesteps - 1), each = length(layers)),
        simplify = FALSE, USE.NAMES = FALSE)),
      atoutput = do.call(c, result3d),
      stringsAsFactors = FALSE)
  }

  if (length(at_data2d) >= 1) {
    # Only select biomasspools if selected variable is "N"!
    if (select_variable == "N") {
      int_fs <- final_species[is.element(final_species, bps)]
      int_fa <- final_agecl[is.element(final_species, bps)]
    }
    # age-structured invert groups are combined in ncdf file!
    if (select_variable == "Grazing") int_fa <- 1
    for (i in seq_along(at_data2d)) {# for loop over all variables!
      if (i == 1) result2d <- list()
      for (j in 1:n_timesteps) {# loop over timesteps
        if (j == 1) values <- array(dim = c(length(boxes), n_timesteps))
        values[, j] <- at_data2d[[i]][, j][boxes + 1]
      }
      result2d[[i]] <- as.vector(values)
    }

    # Order of the data in value column = "atoutput".
    # 1. species  --> rep each with the number of
    #                 ageclasses and n_timesteps * boxes
    # 2. age      --> rep each (1:10 for each species) with n_timesteps * boxes
    # 3. timestep --> rep each timestep (1:n_timesteps)
    #                 with the number of boxes and final_agecl
    #                 (num cohorts per species)
    # 4. polygon  --> rep boxes times n_timesteps * final_agecl
    #                 (num cohorts per species)
    result2d <- data.frame(species = unlist(sapply(
      X = mapply(FUN = rep, x = int_fs, each = int_fa, SIMPLIFY = FALSE,
        USE.NAMES = FALSE),
        FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
      agecl = unlist(sapply(X = sapply(X = int_fa, FUN = seq, from = 1,
        by = 1, simplify = FALSE, USE.NAMES = FALSE),
        FUN = rep, each = length(boxes) * n_timesteps, simplify = FALSE)),
      polygon = unlist(sapply(X = n_timesteps * int_fa,
        FUN = rep, x = boxes, simplify = FALSE, USE.NAMES = FALSE)),
      time = unlist(sapply(X = int_fa, FUN = rep, x = rep(0:(n_timesteps - 1),
        each = length(boxes)), simplify = FALSE, USE.NAMES = FALSE)),
      atoutput = do.call(c, result2d),
      stringsAsFactors = F)
    if (select_variable == "N") result2d$layer <- n_layers - 1
  }

  # Combine dataframes if necessary!
  if (all(sapply(lapply(at_data, dim), length) == 3) & select_variable != "N") {
    result <- result3d
  }
  if (all(sapply(lapply(at_data, dim), length) == 2) & select_variable != "N") {
    result <- result2d
  }
  if (select_variable == "N") {
    if (length(at_data2d) >= 1 & length(at_data3d) == 0) result <- result2d
    if (length(at_data2d) == 0 & length(at_data3d) >= 1) result <- result3d
    if (length(at_data2d) >= 1 & length(at_data3d) >= 1) {
      result <- rbind(result2d, result3d)
    }
  }

  # Remove min_pools if existent (well, there always are min pools... ;)).
  min_pools <- is.element(result$atoutput, c(0, 1e-08, 1e-16))
  if (length(min_pools) > 0) {
    # exclude 1st timestep and sediment layer from calculation
    print_min_pools <- sum(min_pools) -
      length(result[min_pools & result$time == 1, 1]) -
      length(result[min_pools & result$time > 1 & result$layer == 7, 1])
    if (print_min_pools > 0 & verbose) {
      warning(paste0(round(print_min_pools/dim(result)[1] * 100),
        "% of ", select_variable, " are true min-pools (0, 1e-08, 1e-16)"))
    }
    result <- result[!min_pools, ]
  }

  # Remove non-existent layers.
  # WARNING: Biomass is build up (very few) in sediment layer for
  # NON sediment groups (e.g. baleen whales)
  # Therefore, I subset all data from that layer for non biomass groups and
  # groups which cannot penetrate into the sediment!
  # UPDATE: Doesn't work with layers as species are not distributed through
  # the whole water column and do not appear in
  # every polygon.

  # Sum up N for invert cohorts if invert cohorts are present!
  # NOTE: invert cohorts of size 10 are not considered!
  if (select_variable == "N" & any(final_agecl != 10 & final_agecl > 1)) {
    result <- result %>%
      dplyr::group_by(polygon, layer, species, time) %>%
      dplyr::summarise(atoutput = sum(atoutput))
  }

  return(result)
}

Does it work?

```{r testloadnc}

file_nc <- ("outputnordic_runresults_01ANNAGEBIO.nc")
bps <- load_bps(dir = d.name, fgs = functional.groups.file, file_init = biomass.pools.file)
allboxes <- load_box(dir = d.name, file_bgm = box.file)
boxes <- get_boundary(allboxes)

numsage <- load_nc_annage(dir = d.name,
                  file_nc = file_nc,
                  bps = bps,
                  fgs = fgs,
                  biolprm = biolprm,
                  select_groups = select_groups,
                  select_variable = "Nums",
                  check_acronyms = TRUE,
                  bboxes = boxes)
  if(verbose) message("Numbers read in from ANNAGEBIO.")

weightage <- load_nc_annage(dir = d.name,
                  file_nc = file_nc,
                  bps = bps,
                  fgs = fgs,
                  biolprm = biolprm,
                  select_groups = select_groups,
                  select_variable = "Weight",
                  check_acronyms = TRUE,
                  bboxes = boxes)
  if(verbose) message("Weight read in from ANNAGEBIO.")

Need to aggreagate these to see if they look reasonable

```{r fullagecomps-test}

omlist_ss<-NOBAom_cod
source(here("/config/NOBA2Config.R"))
source(here("/config/omdimensions.R"))
source(here("/config/codsurvey.R"))
source(here("/config/codfishery.R"))

n_reps <- 1

  #numbers based fishery independent survey for age and length comps
  # same user specifications as indices
  survey_N <- atlantisom::create_survey(dat = numsage,
                                        time = survtime,
                                        species = select_groups,
                                        boxes = survboxes,
                                        effic = surveffic,
                                        selex = survselex)
  #Sample fish for age composition
  # if we want replicates for obs error this sample function will generate them
  age_comp_data <- list()
  for(i in 1:n_reps){
    age_comp_data[[i]] <- atlantisom::sample_fish(survey_N, surveffN)
  }

    # save age comps
  if(save){
    saveRDS(age_comp_data, file.path(d.name, paste0(scenario.name, "survObsFullAgeComp.rds")))
  }
